Ecosystem size and complexity dictate riverine biodiversity

### 2021-02-23 manuscript.Rmd
Update manuscript.Rmd

---
title: "Ecosystem size and complexity dictate riverine biodiversity"
author:
- Akira Terui^[Department of Biology, University of North Carolina at Greensboro, Greensboro, NC 27412, USA]
- Seoghyun Kim\footnotemark[1]
- Christine L. Dolph^[Department of Ecology, Evolution and Behavior, University of
  Minnesota, St. Paul, MN 55108, USA]
- Taku Kadoya^[National Institute for Environmental Studies, Tsukuba, Ibaraki 305-8506, Japan]
- Yusuke Miyazaki^[Shiraume Gakuen College, Kodaira, Tokyo 187-8570, Japan]
output: word_document
abstract: The species-area relationship, a tendency in which species richness increases with increasing area, is a ubiquitous pattern in ecology[@lomolino2000] and is the foundation of many conservation strategies[@myers2000; @desmet2004; @neigel2003]. However, many ecosystems possess a complex spatial structure that cannot be represented simply by area[@rodríguez-iturbe2001; @turner2015], and the role of such scale-invariant complexity in scaling biodiversity relationships is largely unknown. Here, we use theory and empirical analysis of extensive fish community data to show that ecosystem size and complexity jointly regulate biodiversity in rivers. Rivers are a prime example of fractal branching networks where the recurrent merging of diverse streams generates a pronounced self-similarity in complex branching patterns[@rodríguez-iturbe2001]. Ecosystem complexity, which we define as the probability of branching per unit river distance[@terui2018; @yeakel2014], should increase watershed-scale species richness by increasing habitat heterogeneity. Meanwhile, ecosystem size should dictate metacommunity size and total habitat diversity, two factors that regulate watershed-scale species richness[@hubbell2001; @kallimanis2008]. Metacommunity simulations in branching networks predicted biodiversity patterns consistent with our hypothesis. Larger and more branched ecosystems supported greater species richness at the watershed scale, and the importance of spatial difference in species composition was increasingly apparent when simulated species dispersal was limited. Cross-watershed comparisons across two geographic regions (Hokkaido Island in Japan and the Midwestern US) provided empirical support for the theoretical predictions. In both regions, watershed-scale species richness increased with ecosystem size (watershed area) and complexity (branching probability) through enhanced spatial difference in species composition. Recognizing the dual control of biodiversity should broaden viable options for spatial planning of protected areas or restoration, thereby helping conserve biodiversity from societal demands that threaten it.

>>>>

---
title: "Ecosystem size and complexity dictate riverine biodiversity"
author:
- Akira Terui^[Department of Biology, University of North Carolina at Greensboro, Greensboro, NC 27412, USA]
- Seoghyun Kim\footnotemark[1]
- Christine L. Dolph^[Department of Ecology, Evolution and Behavior, University of
  Minnesota, St. Paul, MN 55108, USA]
- Taku Kadoya^[National Institute for Environmental Studies, Tsukuba, Ibaraki 305-8506, Japan]
- Yusuke Miyazaki^[Shiraume Gakuen College, Kodaira, Tokyo 187-8570, Japan]
output: word_document
abstract: The species-area relationship, a tendency in which species richness increases with increasing area, is a ubiquitous pattern in ecology[@lomolino2000] and is the foundation of many conservation strategies[@myers2000; @desmet2004; @neigel2003]. However, many ecosystems possess a complex spatial structure that cannot be represented simply by area[@rodríguez-iturbe2001; @turner2015], and the role of such scale-invariant complexity in scaling biodiversity relationships is largely unknown. Here, we use theory and empirical analysis of extensive fish community data to show that ecosystem size and complexity jointly regulate biodiversity in rivers. Rivers are a prime example of fractal branching networks where the recurrent merging of diverse streams generates a pronounced self-similarity in complex branching patterns[@rodríguez-iturbe2001]. Ecosystem complexity, which we define as the probability of branching per unit river distance[@terui2018; @yeakel2014], should increase watershed-scale species richness by increasing habitat heterogeneity. Meanwhile, ecosystem size should dictate metacommunity size and total habitat diversity, two factors that regulate watershed-scale species richness[@hubbell2001; @kallimanis2008]. Metacommunity simulations in branching networks predicted biodiversity patterns consistent with our hypothesis. Larger and more branched ecosystems supported greater species richness at the watershed scale, and the importance of local environmental filtering of species was increasingly apparent when simulated species dispersal was limited. Cross-watershed comparisons across two geographic regions (Hokkaido Island in Japan and the Midwestern US) provided empirical support for the theoretical predictions. In both regions, watershed-scale species richness increased with ecosystem size (watershed area) and complexity (branching probability) through enhanced spatial difference in species composition. Recognizing the dual control of biodiversity should broaden viable options for spatial planning of protected areas or restoration, thereby helping conserve biodiversity from societal demands that threaten it.
